---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# read.sql

<!-- <img src='man/figures/' align="right" height="131.5" /></a> -->

<!-- badges: start -->

<!-- badges: end -->

I believe SQL should be written in a seperate file that can be modified later easily instead of a string variable that we all are used to. It also helps you add a proper DB admin into the team who can modify queries without ever learning R. This package consists of few very simple functions that I have been using in my projects for very long and I see myself rewriting them again and again. This package doesn't have any dependencies that is not useful while talking to a DB. I would recommend to use it whenever you have SQL files in your project.

## Installation

It's not on cran yet. The development version can be installed from [GitHub](https://github.com/) with:

```r
# install.packages("devtools")
devtools::install_github("vikram-rawat/read_sql_files")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(read.sql)
## basic example code
```

Always start with creating a config file for keeping all your basic environment variables in a file, specially db configurations. It looks something like this

    default:

      datawarehouse:
        server: localhost
        uid: postgres
        pwd: postgres
        port: 5432
        database: chatbot

Then you can read it with config package. make sure you don't add a drv name in the YML file and also avoid using the password directly into Yml file. Use environment variable instead. read.sql have 1 function to create either a connection or a pool directly from this list.

```{r eval=FALSE}

dw <- config::get("datawarehouse")

conn <- read.sql::rs_create_conn(
  driver = RPostgres::Postgres(),
  param_list = dw
)

pool <- read.sql::rs_create_conn(
  driver = RPostgres::Postgres(),
  param_list = dw,
  pool = TRUE
)

```

These functions come in handy when you want to re-establish a connection. So I preferred to use them in the file.

Then there are only 3 functions that remain

```{r connectdb}

conn <- DBI::dbConnect(RSQLite::SQLite(), ":memory:")

DBI::dbWriteTable(conn,"iris", iris)

```

Always write your SQL code in a separate file where you can edit the code for later use or you can give it somebody who can optimize the code. An external file helps you write and maintain large SQL code. Now imagine you have a code like this.

### get_sql_query

```{sql connection=conn}

select
  *
from
  iris
limit
  5

```

In this case you can use the function like this

    query <- read.sql::rs_read_query(filepath = "path/to/sql/file", method = "DBI::dbGetQuery")

Now suppose you have a query like this.

```{sql connection=conn}

select
  *
from
  iris
where
  `Sepal.Length` > 5
  and
  `Petal.Length` < 1.7

```

what if you want to make this query reuse able and use multiple parameters. You could use SQL interpolations like this.

```{sql eval=FALSE}

select
  *
from
  iris
where
  `Sepal.Length` > ?minsepal
  and
  `Petal.Length` < ?minpetal

```

and then you could use it in the function as this.

```{r eval=FALSE}
sql_query_object <- read.sql::rs_interpolate(
  sql_query = sql_query_object, # object created from rs_read_query function
  sql_conn = conn,
  query_params = list(
    minsepal = 5,
    minpetal = 5
  )
)
```

query_params is an optional argument. You would only need it when you want to interpolate a value in the SQL file you need.

### update meta data in SQL query

sometimes you will feel like you want to dynamically update the table name or column name in a query. This functionality is not available in `DBI::sqlInterpolate` function. Hence this has to be dealt with seperately. So there is an optional parameter called `meta_query_params` in the function `rs_interpolate` which will simply replace anything inside the `{}` curly brackets by matching it with named list. liket this

```{sql eval=FALSE}

select
  *
from
  {main_table}
where
  {column1} >= ?mincol1
  and
  {column2} >= ?mincol2
  and
  {column3} in ({species_value})
  and
  `Petal.Width` in ({width_value})

```

to run a query like this we need to pass 2 values like this.

```{r eval=FALSE}

query_obj <- read.sql::rs_interpolate(
  sql_query = sql_query_object, # object created from rs_read_query function
  sql_conn = conn,
  meta_query_params = list(
    main_table = "iris",
    column1 = "`Sepal.Length`",
    column2 = "`Petal.Length`",
    column3 = "`Species`",
    species_value = c("Setosa", "versicolor"),
    width_value = seq(1.0, 1.4, 0.1)
  ),
  query_params = list(
    mincol1 = 4,
    mincol2 = 4
  )
)

```

### execute_sql_file

The rs_execute function is the core execution engine. It takes the final sql_query object (which contains the prepared SQL and the execution method) and runs it against the database connection.

This function is highly flexible and automatically handles argument passing for the most common database packages, simplifying your execution calls.

#### Execution Method (method argument of rs_read_query)

The method used for execution is specified when you call rs_read_query. rs_execute uses the following rules to determine which arguments to pass:

##### 1. DBI Methods (Reading or Writing):

If the method is DBI::dbGetQuery or DBI::dbExecute, rs_execute automatically passes the arguments as (conn, statement, ...).

##### 2. ADBC Methods (Reading or Writing):

If the method is read_adbc or execute_adbc, rs_execute automatically passes the arguments as (db_or_con, query, ...).

##### 3. Custom/Other Methods:

If the method is any other function (e.g., dbSendStatement, a custom S7 method, etc.), rs_execute requires the user to manually pass the connection object via the optional ... arguments. It will only automatically pass the SQL statement using the argument name specified by stmt_arg_name (defaulting to "statement").

```{r eval=FALSE}
# 1. DBI (Default Read) - Automatic Argument Handling
query_obj_read <- read.sql::rs_read_query(
    filepath = "path/to/read.sql",
    method = "DBI::dbGetQuery"
)

query_obj_read |>
    read.sql::rs_execute(sql_conn = conn)
# Executes: DBI::dbGetQuery(conn = conn, statement = SQL)

# 2. ADBC (Write/Execute) - Automatic Argument Handling
query_obj_post <- read.sql::rs_read_query(
    filepath = "path/to/update.sql",
    method = "adbc::execute_adbc"
)

query_obj_post |>
    read.sql::rs_execute(sql_conn = conn)
# Executes: adbc::execute_adbc(db_or_con = conn, query = SQL)

# 3. Custom Method (Requires Manual Connection Passing via ...)
query_obj_custom <- read.sql::rs_read_query(
    filepath = "path/to/custom.sql",
    method = "my_package::custom_exec_fn"
)

query_obj_custom |>
    read.sql::rs_execute(
        sql_conn = conn,
        conn_arg = conn # User must supply the connection here
    )
# Executes: my_package::custom_exec_fn(conn_arg = conn, statement = SQL)
```

The only thing different about it is that it has a method argument where if you need results from the DB you should use `get` all lower case. If you want to execute a delete, or update statement you should use `post` all lower case.

### migration

The rs_migrate function safely applies a series of SQL scripts saved in the sql/migrate/up or sql/migrate/down folders.

Crucially, this function wraps the entire sequence of script executions in a single database transaction.

If all scripts run successfully, the transaction is committed.

If any script fails, the transaction is automatically rolled back, ensuring the database is left in its original, consistent state.

```{r eval=FALSE}
# Execute all scripts in the 'up' folder within a single transaction
conn |>
  rs_migrate(
    up = TRUE,
    default_method = "DBI::dbExecute"
  )

```

The function relies on the DBI::dbWithTransaction generic, so the connection object supplied must support DBI transactions (e.g., standard RPostgres, RSQLite, etc., connections). This will help you set up a DB again and again anytime you need it.

### warning

Minor refinement to emphasize the crashing behavior is intentional for reliable execution.

Package doesn't assume anything and it does no checking at all. It is meant to be used with existing architecture where you will write all the logic on top of it. So if there is anything wrong it will simply crash.

Please read the documentation of each function to understand different arguments used in them.
